---
title: "Tomato Bag Exploratory Data Analysis - Orders by State"
author: "Karl Nickl - knickl@uchicago.edu"
date: "January 26, 2019"
output: html_document
---

## Prepare the Environment
### Load required libraries
```{R}
library(scales)  # for graph formatting
library(tidyverse)  # for data manipulation and plotting
library(gridExtra)  # for arranging plots
library(grid)  # for arranging plots
library(showtext)
library(directlabels)
```

```{r}
# add the Roboto font from Google Fonts
font_add_google("Roboto", "roboto")

# turn on showtext font usage
showtext_auto(enable = TRUE)
```

### Load project utilities
```{r}
source("../../capstone_engineering/data_prep_utils.R")
```

## Load and Prepare Data
### Read the data
```{r}
# Path for cleaned sales data
sales_path <- "../../capstone_engineering/data/clean/Sales_data_for_Capstone_Project_-_Student_Version_-_v3_CLEAN.csv"

# Read the sales data
schema = cols(
  fiscalquartername                 = col_character(),
  FiscalPeriod                      = col_integer(),
  FiscalYear                        = col_integer(),
  Company                           = col_character(),
  Top_Most_BP                       = col_character(),
  BusinessPartner                   = col_character(),
  Item_Number                       = col_character(),
  Item_Description                  = col_character(),
  Ship_to_State                     = col_character(),
  Ship_to_City                      = col_character(),
  Quantity                          = col_number(),
  Planned_Delivery_Date             = col_datetime(format = "%Y-%m-%dT%H:%M:%SZ"),
  Address_1                         = col_character(),
  Address_2                         = col_character(),
  Postal_Code                       = col_character(),
  Budget_Qty                        = col_integer(),
  Forecast_Qty                      = col_integer(),
  Line_of_Business                  = col_character(),
  Ship_to_State_clean               = col_factor(levels = NULL),
  is_domestic                       = col_logical(),
  planned_delivery_year             = col_integer(),
  planned_delivery_month_abbr       = col_factor(
    levels  = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
    ordered = TRUE  
  ),
  planned_delivery_month_num        = col_integer(),
  planned_delivery_day_of_week_abbr = col_factor(
    levels  = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"),
    ordered = TRUE
  ),
  planned_delivery_day_of_week_num  = col_integer()
)

sales_dat <- read_csv(
  file      = sales_path,
  col_types = schema,
  na        = c("", "NA", "NULL")
)

# Read the integrated sales data  ----
sales_dat <- sales_dat %>% 
  mutate(state_is_california = case_when(Ship_to_State == "CA" ~ "CA",
                                         Ship_to_State != "CA" ~ "Other",
                                         is.na(Ship_to_State)  ~ "Other"),
         is_ca_or_dr         = case_when(
           Ship_to_State %in% c("CA", "DR") ~ "yes",
           TRUE ~ "no"
           ),
         Ship_to_State       = Ship_to_State_clean
      )
```

### Prepare aggregated data for plotting

#### Count of States by Year
```{r}
states_by_year_dat <- sales_dat %>%
  group_by(planned_delivery_year) %>%
  summarise(state_count = n_distinct(Ship_to_State))
  
head(states_by_year_dat)
```

#### Total Orders by State
Also extract top states based on order quantity
```{r}
orders_by_state <- sales_dat %>%
  group_by(Ship_to_State) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )

top_states <- orders_by_state %>%
  top_n(5, total_quantity) %>%
  pull(Ship_to_State)

top_state <- orders_by_state %>%
  top_n(1, total_quantity) %>%
  pull(Ship_to_State)



# classify top states
sales_dat <- sales_dat %>%
  mutate(
    top_state_class = case_when(
      Ship_to_State == top_state    ~ as.character(Ship_to_State),
      Ship_to_State %in% top_states ~ "Top 5 States (w/out CA)",
      TRUE                          ~ "Other"
    ),
    graphing_state = case_when(
      Ship_to_State %in% top_states ~ as.character(Ship_to_State),
      TRUE                          ~ "Other"
    )
  )

orders_by_city <- sales_dat %>%
  group_by(Ship_to_City) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )

top_cities <- orders_by_city %>%
  top_n(5, total_quantity) %>%
  pull(Ship_to_City)

top_city <- orders_by_city %>%
  top_n(1, total_quantity) %>%
  pull(Ship_to_City)

sales_dat <- sales_dat %>%
  mutate(
    top_city_class = case_when(
      Ship_to_City %in% top_cities ~ "Top 5 Cities",
      TRUE                          ~ "Other"
    ),
    graphing_city = case_when(
      Ship_to_City %in% top_cities ~ as.character(Ship_to_City),
      TRUE                          ~ "Other"
    )
  )
  
# sales data filtered to top states
top_sales_dat <- sales_dat %>% filter(Ship_to_State %in% top_states)
top_sales_dat$Ship_to_State <- droplevels(top_sales_dat$Ship_to_State)

# sales data filtered to top states (less California)
sales_no_ca_dat <- top_sales_dat %>% filter(Ship_to_State != "CA")

orders_by_ca_dat <- sales_dat %>%
  group_by(state_is_california) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
```

#### Orders by State and Year
```{r}
orders_by_state_by_year_dat <- sales_dat %>%
  group_by(Ship_to_State, planned_delivery_year) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
head(orders_by_state_by_year_dat)
```

#### Orders by State and Year 
```{r}
orders_by_ca_by_year_dat <- sales_dat %>%
  group_by(state_is_california, planned_delivery_year) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
orders_by_year_dat <- sales_dat %>% 
  group_by(planned_delivery_year) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  ) %>%
  mutate(state_is_california = "Total")
orders_by_ca_by_year_dat <- orders_by_ca_by_year_dat %>% 
  union_all(orders_by_year_dat)
head(orders_by_ca_by_year_dat)
```

#### Orders by State by Month
```{r}
# Create a dataframe of orders by state by month
orders_by_state_by_month_dat <- top_sales_dat %>%
  group_by(Ship_to_State, planned_delivery_month) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
head(orders_by_state_by_month_dat)
```

#### Orders by California by Month
```{r}
# Create a dataframe of orders by state by month
orders_by_ca_by_month_dat <- sales_dat %>%
  group_by(state_is_california, planned_delivery_month) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
head(orders_by_state_by_month_dat)
```

#### Order Quantity by State and Fulfilling Plant
```{r}
orders_by_state_and_plant <- top_sales_dat %>%
  group_by(Ship_to_State, Company) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
head(orders_by_state_and_plant)
```

#### Order Quantity by California and Fulfilling Plant
```{r}
orders_by_state_and_plant <- top_sales_dat %>%
  group_by(state_is_california, Company) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )
head(orders_by_state_and_plant)
```

## Exploratory Plots
#### Order Quantity by California
```{r, dpi=400, fig.width=6.7, fig.height=4}
plt <- orders_by_ca_dat %>%
  ggplot(aes(
    x    = reorder(state_is_california, total_quantity),
    y    = total_quantity,
    fill = reorder(state_is_california, total_quantity))
  ) +
  geom_bar(stat = "identity", width = .4) +
  scale_y_continuous(
    label  = comma, 
    name   = "Total Order Quantity", 
    expand = c(0, 0),
    limits = c(-100, 17000000)
  ) +
  scale_x_discrete(name = element_blank()) +
  ggtitle(
    "California drives tomato bag sales",
    "Total tomato bag order quantity, 2010-2019"
  ) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.x       = element_text(face = "bold", margin = margin(t = 10)),
    axis.text.y        = element_text(face = "bold", margin = margin(r = 8)),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,18,18,13), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.x = element_line(color = "#E6E6E6", size = 1),
    legend.position    = "none"
  ) +
  scale_fill_manual(values = c("CA" = "#F6303E", "Other" = "#2A373D")) +
  coord_flip()
plt

```

#### All States Order Quantity
```{r, dpi=400, fig.width=6.7, fig.height=4}
orders_by_state %>%
  filter(Ship_to_State != "CA") %>%
  mutate(is_top_3_qty = if_else(rank(-total_quantity) <= 4, "top", "not")) %>%
  ggplot(aes(
    x = reorder(Ship_to_State, -total_quantity), 
    y = total_quantity, 
    fill = is_top_3_qty
  )) +
  geom_bar(stat = "identity", width = .4) +
  scale_y_continuous(
    label  = comma, 
    name   = "Total Order Quantity", 
    expand = c(0, 0),
    limits = c(-400, 500000)
  ) +
  scale_x_discrete(name = element_blank()) +
  ggtitle(
    "The remainder of orders are driven by four states/provinces",
    "Total tomato bag order quantity with California omitted, 2010-2019"
  ) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(margin = margin(t = 10)),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,18,13,13), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position    = "none"
  ) +
  scale_fill_manual(values = c("top" = "#F6303E", "not" = "#2A373D")) 
  
```
#### Annual Order Quantity by CA
```{r, dpi=400, fig.width=6.7, fig.height=4}
plt <- orders_by_ca_by_year_dat %>% 
  ggplot(aes(
    x        = planned_delivery_year, 
    y        = total_quantity,
    color    = state_is_california,
    linetype = state_is_california
  )) +
  geom_line(size = 1.3) +
  scale_y_continuous(
    label  = comma, 
    name   = "Annual Order Quantity", 
    expand = c(0, 0),
    limits = c(0, 2250000)
  ) +
  scale_x_continuous(name = element_blank()) +
  ggtitle(
    "Annual sales demonstrate no visible trend",
    "Annual tomato bag order quantity, 2010-2019"
  ) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(margin = margin(t = 10)),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,18,13,13), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.title       = element_blank(),
    legend.position    =  c(0.85, 0.4)
  ) +
  scale_color_manual(values = c(
    "CA"         = "#2A373D",
    "Total"      = "#F6303E",
    "Other" = "#2A373D"
  )) +
  scale_linetype_manual(values = c(
    "CA"         = 2, 
    "Total"      = 1, 
    "Other" = 1
  ))
plt
```

#### Annual sales (location)/no cali
```{r, fig.width=6.7, fig.height=4, dpi=400}
axis_min <- orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  group_by(1) %>%
  summarise(min_val = min(total_quantity)) %>%
  pull()
axis_max <- orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  group_by(1) %>%
  summarise(max_val = max(total_quantity)) %>%
  pull()

orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = total_quantity
  )) +
  geom_line(size = 1.3, color = "#F6303E") +
  scale_y_continuous(
    label  = comma, 
    name   = "Annual Order Quantity", 
    expand = c(0,0),
    limits = c(axis_min, axis_max)
  ) +
  scale_x_continuous(name = element_blank()) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_blank(),
    axis.title.x       = element_blank(),
    axis.text.x        = element_text(margin = margin(t = 7), size = 6),
    axis.text.y        = element_text(margin = margin(t = 7), size = 6),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,18,13,19), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = .8),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "#E6E6E6", size = .8)
  ) +
  facet_wrap(~ Ship_to_State, ncol = 4) +
  ggtitle("Annual demand varies annually in Ohio, Ontario, and Arizona",
          "Annual tomato bag order quantity by location with California omitted, 2010-2018")
  
```

#### Distribution of Order Size
```{r, dpi=500, fig.width=6.7, fig.height=4}
mean_order_size <- mean(sales_dat$Quantity)

sales_dat %>% 
  ggplot(aes(x = Quantity)) +
  geom_histogram(color = "#FFFFFF", fill = "#2A373D", bins = 10) +
  scale_y_continuous(
    label  = comma, 
    name   = "Frequency", 
    expand = c(0, 0),
    limits = c(0, 400)
  ) +
  scale_x_continuous(label = comma, name = "Order Size") +
  ggtitle(
    "Order size is right skewed",
    "Distribution of tomato bag order quantity, 2010-2019"
  ) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(margin = margin(t = 10)),
    axis.title.x       = element_text(face = "bold"),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,18,13,13), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_color_manual(values = c(
    "CA"         = "#2A373D",
    "Total"      = "#F6303E",
    "Other" = "#2A373D"
  )) +
  scale_linetype_manual(values = c(
    "CA"         = 2, 
    "Total"      = 1, 
    "Other" = 1
  ))
```
#### Distribution by State
```{r, dpi=500, fig.width=6.7, fig.height=4}
sales_dat %>%
  filter(!is.na(Ship_to_State)) %>%
  ggplot(aes(x = Ship_to_State, y = Quantity, color = is_ca_or_dr)) +
  geom_jitter(size = .5, color = "#333F49", height = 0, width = .2) +
  geom_boxplot(fill = "#FFFFFF", outlier.shape = NA, size = .8)  +
  scale_y_continuous(label = comma, name = "Order Size") +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(margin = margin(t = 10)),
    axis.title.x       = element_blank(),
    plot.title         = element_text(face = "bold"),
    plot.margin        = unit(c(15,12,13,13), "pt"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    legend.position    = "none"
  ) +
  scale_color_manual(values = c("yes" = "#F6303E", "no" = "#2A373D")) +
  scale_x_discrete() +
  ggtitle(
    "California & the Dominican republic have large orders",
    paste0("Distribution of tomato bag order quantity by location, 2010-2018 (",
           "jitter added)")
  )
  

```



#### Monthly Order Quantity
```{r, dpi=500, fig.width=6.7, fig.height=4}
orders_by_year_month_dat <- sales_dat %>% 
  group_by(planned_delivery_year, planned_delivery_month) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )

monthly_order_summary <- orders_by_year_month_dat %>%
  group_by(planned_delivery_month) %>%
  summarise(
    total_orders_by_mo      = sum(total_orders),
    total_qty_by_mo         = sum(total_quantity),
    mean_monthly_quantity   = mean(total_quantity),
    median_monthly_quantity = median(total_quantity),
    max_monthly_quantity    = max(total_quantity),
    min_monthly_quantity    = min(total_quantity),
    sd_monthly_quantity     = sd(total_quantity, na.rm = TRUE)
  )

monthly_order_summary %>%
  ggplot(aes(x = planned_delivery_month)) +
  geom_ribbon(
    aes(
      x    = 1:length(planned_delivery_month),
      ymin = mean_monthly_quantity - 2 * sd_monthly_quantity,
      ymax = mean_monthly_quantity + 2 * sd_monthly_quantity
    ),
    alpha = .5,
    fill = "#333F49"
  ) +
  geom_line(
    aes(y = mean_monthly_quantity, x = planned_delivery_month),
    color = "#F6303E",
    size = 1.3,
    group = 1) +
  scale_y_continuous(label = comma, name = "Monthly Order Quantity") +
  scale_x_discrete(labels = month.abb) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(),
    axis.title.x       = element_blank(),
    plot.title         = element_text(face = "bold"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    plot.margin        = unit(c(15,12,13,13), "pt")
  ) +
  ggtitle(
    "Bag demand peaks during the Summer",
    "Mean monthly tomato bag order quantity +/- 2 standard deviations, 2010-2018"
  )
  
  
```

#### Same as above but for CA
```{r, r, dpi=500, fig.width=6.7, fig.height=4}

orders_by_year_month_ca_dat <- sales_dat %>%
  filter(Ship_to_State == "CA") %>%
  group_by(planned_delivery_year, planned_delivery_month) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )

monthly_order_ca_summary <- orders_by_year_month_dat %>%
  group_by(planned_delivery_month) %>%
  summarise(
    total_orders_by_mo      = sum(total_orders),
    total_qty_by_mo         = sum(total_quantity),
    mean_monthly_quantity   = mean(total_quantity),
    median_monthly_quantity = median(total_quantity),
    max_monthly_quantity    = max(total_quantity),
    min_monthly_quantity    = min(total_quantity),
    sd_monthly_quantity     = sd(total_quantity, na.rm = TRUE)
  )



monthly_order_ca_summary %>%
  ggplot(aes(x = planned_delivery_month)) +
  geom_ribbon(
    aes(
      x    = 1:length(planned_delivery_month),
      ymin = mean_monthly_quantity - 2 * sd_monthly_quantity,
      ymax = mean_monthly_quantity + 2 * sd_monthly_quantity
    ),
    alpha = .5,
    fill = "#333F49"
  ) +
  geom_line(
    aes(y = mean_monthly_quantity, x = planned_delivery_month),
    color = "#F6303E",
    size = 1.3,
    group = 1) +
  scale_y_continuous(label = comma, name = "Monthly Order Quantity") +
  scale_x_discrete(labels = month.abb) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(),
    axis.title.x       = element_blank(),
    plot.title         = element_text(face = "bold"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    plot.margin        = unit(c(15,12,13,13), "pt")
  ) +
  ggtitle(
    "Bag demand peaks during the Summer in California",
    "Mean California mo. tomato bag order qty. +/- 2 standard deviations, 2010-2018"
  )
```



#### above by state (no CA)
```{r, fig.width=6.7, fig.height=4, dpi=400}
orders_by_year_month_state_dat <- sales_dat %>% 
  filter(Ship_to_State != "CA") %>%
  group_by(planned_delivery_year, planned_delivery_month, Ship_to_State) %>%
  summarise(
    total_orders        = n(),
    total_quantity      = sum(Quantity),
    mean_order_quantity = mean(Quantity),
    sd_order_quantity   = sd(Quantity)
  )

monthly_state_order_summary <- orders_by_year_month_state_dat %>%
  group_by(planned_delivery_month, Ship_to_State) %>%
  summarise(
    total_orders_by_mo      = sum(total_orders),
    total_qty_by_mo         = sum(total_quantity),
    mean_monthly_quantity   = mean(total_quantity),
    median_monthly_quantity = median(total_quantity),
    max_monthly_quantity    = max(total_quantity),
    min_monthly_quantity    = min(total_quantity),
    sd_monthly_quantity     = sd(total_quantity, na.rm = TRUE)
  )

months_states_dat <- expand.grid(
  monthly_state_order_summary %>% 
    select(Ship_to_State) %>% pull() %>% unique(),
  monthly_state_order_summary %>% 
    select(planned_delivery_month) %>% distinct() %>% pull()
) %>% distinct()

colnames(months_states_dat) <- c("Ship_to_State", "planned_delivery_month")

monthly_state_order_summary <- monthly_state_order_summary %>%
  full_join(months_states_dat)

monthly_state_order_summary <- monthly_state_order_summary %>%
  mutate(
    mean_monthly_quantity = case_when(
      is.na(mean_monthly_quantity) ~ 0,
      TRUE                         ~ mean_monthly_quantity
    ),
    sd_monthly_quantity   = case_when(
      is.na(sd_monthly_quantity) | 
        is.nan(sd_monthly_quantity) ~ 0,
      TRUE                          ~ sd_monthly_quantity
    ))

monthly_state_order_summary$month_num <- match(
  monthly_state_order_summary$planned_delivery_month,
  month.abb)

min_qty <- min(monthly_state_order_summary$min_monthly_quantity) * 1.05
max_qty <- max(monthly_state_order_summary$max_monthly_quantity) * 1.05

monthly_state_order_summary %>%
  ggplot(aes(x = planned_delivery_month)) +
  geom_ribbon(
    aes(
      x    = month_num,
      ymin = mean_monthly_quantity - 2 * sd_monthly_quantity,
      ymax = mean_monthly_quantity + 2 * sd_monthly_quantity
    ),
    alpha = .5,
    fill = "#333F49"
  ) +
  geom_line(
    aes(y = mean_monthly_quantity, x = planned_delivery_month),
    color = "#F6303E",
    size = 1.3,
    group = 1) +
  scale_y_continuous(label = comma, name = "Mean Order Size") +
  scale_x_discrete(breaks = c("Jan", "Mar", "May", "Jul", "Sep", "Nov")) +
  theme_linedraw() +
  theme(
    plot.background    = element_rect(fill = "#E6E6E6"),
    panel.background   = element_rect(fill = "#FFFFFF"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    text               = element_text(color = "#333F49", family = "roboto"),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(size = 6, margin = margin(t = 7)),
    axis.title.x       = element_blank(),
    plot.title         = element_text(face = "bold"),
    panel.border       = element_blank(),
    panel.grid.major.y = element_line(color = "#E6E6E6", size = 1),
    panel.grid.minor.y = element_line(color = "#E6E6E6", size = 1),
    plot.margin        = unit(c(15,18,13,19), "pt")
  ) +
  facet_wrap(~ Ship_to_State, ncol = 4) +
  ggtitle(
    "Bag demand peaks during the Summer",
    "Mean monthly tomato bag order quantity +/- 2 standard deviations, 2010-2018"
  )

```

#### Number of Ship-To States per Year
```{r}
# plot the number of states per year
states_by_year_dat %>%
  ggplot(aes(x = planned_delivery_year, y = state_count)) +
  geom_col() +
  scale_y_continuous(
    label = comma, 
    name = "Number of Locations", 
    expand = c(0, 0)
  ) +
  scale_x_continuous(name = "Planned Delivery Year") +
  ggtitle("Number of Ship to States by Year") +
  theme_light()
```

#### Total Order Quantity by State and Year
```{r}
orders_by_state_by_year_dat %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = total_quantity, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(label = comma, name = "Total Quantity", expand = c(0, 0)) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State") +
  ggtitle("Total Order Quantity by State and Year") +
  theme_light()
```

#### Total Order Quantity by State and Year (CA)
```{r}
orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = total_quantity, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(label = comma, name = "Total Quantity", expand = c(0, 0)) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State") +
  ggtitle("Total Order Quantity by State and Year (w/out California") +
  theme_light()
```

#### Total Order Count by State and Year
```{r}
orders_by_state_by_year_dat %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = total_orders, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
    label = comma, 
    name = "Total Number of Orders", 
    expand = c(0, 0)
  ) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State") +
  ggtitle("Total Number of Orders by State and Year") +
  theme_light()
```

#### Total Order Count by State and Year (w/out CA)
```{r}
orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = total_orders, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
    label = comma, 
    name = "Total Number of Orders", 
    expand = c(0, 0)
  ) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State") +
  ggtitle("Total Number of Orders by State and Year (w/out California)") +
  theme_light()
```

#### Average Order Size by State and Year
```{r}
orders_by_state_by_year_dat %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = mean_order_quantity, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
    label = comma, 
    name = "Mean Order Quantity", 
    expand = c(0, 0)
  ) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State", drop = TRUE) +
  ggtitle("Mean Order Quantity by State and Year") +
  theme_light()
```

#### Average Order Size by State and Year (w/out CA)
```{r}
orders_by_state_by_year_dat %>%
  filter(Ship_to_State != "CA") %>%
  ggplot(aes(
    x = planned_delivery_year, 
    y = mean_order_quantity, 
    fill = Ship_to_State
  )) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
    label = comma, 
    name = "Mean Order Quantity", 
    expand = c(0, 0)
  ) +
  scale_x_continuous(name = "Planned Delivery Year") +
  scale_fill_discrete(name = "Ship to State", drop = FALSE) +
  ggtitle("Mean Order Quantity by State and Year (w/out CA") +
  theme_light()
```



#### Monthly Order Quantity by State
Define a function for the plot
```{r}
plot_order_quantity_by_month_for_state <- function(state, dat) {
  dat <- dat %>% filter(Ship_to_State == state)
  
  plt <- dat %>%
    ggplot(aes(x = planned_delivery_month, y = total_quantity)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(
      label = comma, 
      name = "Total Order Quantity", 
      expand = c(0, 0)
    ) +
    scale_x_discrete(name = "Planned Delivery Month", drop = FALSE) +
    ggtitle(state)
  
  plt
}
```

Create a vector of unique states
```{r}
unique_states <- sales_dat %>%
  pull(Ship_to_State) %>%
  as.character() %>%
  unique() %>%
  sort()
```

Create the plots and arrange
```{r, fig.width=15, fig.height=10}
# Create a list of plot objects
state_monthly_quantity_plt_lst <- lapply(
  unique_states, 
  plot_order_quantity_by_month_for_state,
  dat = orders_by_state_by_month_dat
)
  
# Arrange and print the plots
grid.arrange(
  grobs = state_monthly_quantity_plt_lst, 
  top = textGrob("Total Order Quantity by Month", gp=gpar(fontsize=20))
)
```

#### Total Quantity by State and Fulfilling Plant
```{r}
# orders_by_state_and_plant %>%
#   ggplot(aes(x = Ship_to_State, y = total_quantity, fill = Company)) +
#   geom_bar(position = "dodge", stat = "identity") +
#   scale_y_continuous(
#     name = "Total Order Quantity",
#     label = comma,
#     expand = c(0, 0)
#   ) +
#   ggtitle("Total Quantity by State and Fulfilling Plant") +
#   theme_light()
```

```{r, dpi = 400}
# Plots required for presentatoin

# Total Sales by CA and others
sales_dat %>% 
  group_by(graphing_state) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  arrange(desc(total_quantity)) %>%
  ggplot(aes(x = reorder(graphing_state, -total_quantity), y = total_quantity)) +
  geom_col() +
  theme_light() +
  theme(
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(
    name   = "Total Order Quantity", 
    label  = comma, 
    expand = c(0, 0),
    limits = c(0, 20000000)
  )
  

```

```{r}
top_5_states_by_year <- sales_dat %>%
  group_by(Ship_to_State, planned_delivery_year) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  group_by(planned_delivery_year) %>%
  top_n(n = 5, wt = total_quantity) %>%
  mutate(graphing_state_year = Ship_to_State) %>%
  select(Ship_to_State, planned_delivery_year, graphing_state_year)

sales_dat <- sales_dat %>%
  left_join(
    top_5_states_by_year,
    by = c("Ship_to_State" = "Ship_to_State", 
           "planned_delivery_year" = "planned_delivery_year")
  ) %>%
  mutate(graphing_state_year = if_else(
    is.na(graphing_state_year), 
    "Other", 
    as.character(graphing_state_year)
  ))

sales_dat %>%
  group_by(graphing_state_year, planned_delivery_year) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  group_by(planned_delivery_year) %>%
  top_n(n = 5, wt = total_quantity) %>%
  ggplot(aes(
    x    = planned_delivery_year, 
    y    = total_quantity,
    fill = graphing_state_year, -total_quantity)
  ) +
  geom_col(position = "stack") +
  theme_light() +
  theme(
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(
    name   = "Total Order Quantity", 
    label  = comma, 
    expand = c(0, 0)
  ) +
  labs(fill = "State")
```

```{r}
sales_dat %>% 
  group_by(graphing_city) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  arrange(desc(total_quantity)) %>%
  ggplot(aes(x = reorder(graphing_city, -total_quantity), y = total_quantity)) +
  geom_col() +
  theme_light() +
  theme(
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(
    name   = "Total Order Quantity", 
    label  = comma, 
    expand = c(0, 0)
  )
```

```{r}
top_5_cities_by_year <- sales_dat %>%
  group_by(Ship_to_City, planned_delivery_year) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  group_by(planned_delivery_year) %>%
  top_n(n = 5, wt = total_quantity) %>%
  mutate(graphing_city_year = Ship_to_City) %>%
  select(Ship_to_City, planned_delivery_year, graphing_city_year)

sales_dat <- sales_dat %>%
  left_join(
    top_5_cities_by_year,
    by = c("Ship_to_City" = "Ship_to_City", 
           "planned_delivery_year" = "planned_delivery_year")
  ) %>%
  mutate(graphing_city_year = if_else(
    is.na(graphing_city_year), 
    "Other", 
    as.character(graphing_city_year)
  ))

sales_dat %>% 
  group_by(graphing_city_year, planned_delivery_year) %>%
  summarise(total_quantity = sum(Quantity)) %>%
  group_by(planned_delivery_year) %>%
  top_n(n = 5, wt = total_quantity) %>%
  ggplot(aes(
    x    = planned_delivery_year, 
    y    = total_quantity,
    fill = graphing_city_year
  )) +
  geom_col(position = "stack") +
  theme_light() +
  theme(
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(
    name   = "Total Order Quantity", 
    label  = comma, 
    expand = c(0, 0)
  ) +
  labs(fill = "City")
```






